---
title: "Getting Started with Binomial"
author: "Jane Zhang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE, message= FALSE}
knitr::opts_chunk$set(collapse = T, comment="#>")
library(binomial)
```

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Binomial  

The package `"binomial"` is a minimal implementation for relative caculation of a binomial variable with given trials, probability, successes, and to visualize the relative probabilities in such a series of successes.

## 1) The choosing factor of binomial variable  
You can get the choosing factor with the function `bin_choose(n,k)`, where n represents the total number of trials, and k represents the number of successes of n trials.
```{r}
bin_choose(n=6,k=2)


```

## 2) Get probability of a binomial variable with given trials, success(es) and probability.  

You can get a (or a list of) probability with valid possible number of successes by using the function `bin_probability(success, tirlas, prob)`

```{r}
bin_probability(2,6,0.8)

```

## 3) Get distribution of a binomial variable with given trials and probability.  
You can get a dataframe with probabilities of all possible number of successes by using function `bin_distribution(trials,prob)`  

```{r}
bin_distri <- bin_distribution(trials = 5, prob = 0.5)
bin_distri
```

## 4) Plot of distribution of probabilities.
You can plot the histogram of distribution by using `plot()`
```{r}
library(ggplot2)
plot(bin_distri)

```

## 5) Getting cumulative probabilities for all possible successes values.
You can get a dataframe with probabilities and cumulative probabilites of all possible successes values with function `bin_cumulative(trials,prob)`

```{r}
bin_cu <- bin_cumulative(5,0.5)
bin_cu
```
## 6) Plot cumulative probability distribution

After creating the dataframe object, you can use plot() to get a histogram to visualize the cumulative probability.
```{r}
plot(bin_cu)

```


## 7) Create a `binomial variable`  

You can create a`"binvar` object with the function `bin_variable(trials,prob)`:
```{r}
bin_var <- bin_variable(trials=5,prob=0.5)
bin_var
```


## 8) Get the summary of the binomial variable

After you created a `'binvar'` object,  you can use the function `summary()`
to get a summary of all related measures of this binomial variable

```{r}
binsum1 <- summary(bin_var)
binsum1
```



## 9) More funcitons of measures
You can use the following functions to get the corresponding measures of the binomial variable with valid inputs, trials and probability. 


Measures       | Functions to Use
---------------|---------------------------
mean           | bin_mean(trials,prob)
variance       | bin_variance(trials,prob)
mode           | bin_mode(trials,prob)
skewness       | bin_skewness(trials,prob)  
kurtosis       | bin_kurtosis(trials,prob)



```{r}
bin_mean(5,0.5)
bin_variance(5,0.5)
bin_mode(5,0.5)
bin_skewness(5,0.5)
bin_kurtosis(5,0.5)
```
